---
title: "01_preprocess"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(ggfortify)
source(here("/src/utils.R"))
```


read S1 files
```{r}

s1_rna_features <- read_delim(here("../data/sub1/sc1_Phase1_GE_FeatureMatrix.tsv"), delim="\t")
s1_outcome <- read_delim(here("../data/sub1/sc1_Phase1_GE_Outcome.tsv"), delim="\t")
s1_phenotype <- read_delim(here("../data/sub1/sc1_Phase1_GE_Phenotype.tsv"), delim="\t")

```

read S2 files
```{r}

s2_CN_features <- read_delim(here("../data/sub2/sc2_Phase1_CN_FeatureMatrix.tsv"), delim="\t")
s2_outcome <- read_delim(here("../data/sub2/sc2_Phase1_CN_Outcome.tsv"), delim="\t")
s2_phenotype <- read_delim(here("../data/sub2/sc2_Phase1_CN_Phenotype.tsv"), delim="\t")

```

read S3 files
```{r}

s3_features <- read_delim(here("../data/sub3/sc3_Phase1_CN_GE_FeatureMatrix.tsv"), delim="\t")
s3_outcome <- read_delim(here("../data/sub3/sc3_Phase1_CN_GE_Outcome.tsv"), delim="\t")
s3_phenotype <- read_delim(here("../data/sub3/sc3_Phase1_CN_GE_Phenotype.tsv"), delim="\t")

```


PCA for RNASeq
```{r}
autoplot(prcomp(s1_rna_features[,-1]))
autoplot(prcomp(s1_rna_features[,-1]), data = s1_phenotype, colour = 'CANCER_TYPE')
autoplot(prcomp(s1_rna_features[,-1]), data = s1_phenotype, colour = 'SEX')
autoplot(prcomp(s1_rna_features[,-1]), data = s1_phenotype, colour = 'RACE')
autoplot(prcomp(s1_rna_features[,-1]), data = s1_phenotype, colour = 'WHO_GRADING')
```


how much overlap between the 3 sets
```{r}
dim(s1_outcome)
dim(s2_outcome)
dim(s3_outcome)

length(intersect(s1_outcome$PATIENTID, s2_outcome$PATIENTID))

length(intersect(s1_outcome$PATIENTID, s3_outcome$PATIENTID))
length(intersect(s2_outcome$PATIENTID, s3_outcome$PATIENTID))
```

There is only one patient with UNCLASSIFIED, we will group it with UNKNOWN and change it to NA
```{r}
s1_phenotype <- modify_phenotypes(s1_phenotype)
s2_phenotype <- modify_phenotypes(s2_phenotype)
s3_phenotype <- modify_phenotypes(s3_phenotype)
```

merge data frames
```{r}
s1_features_with_outcome <- inner_join (s1_rna_features, s1_outcome, by="PATIENTID")
s1_features_with_outcome_phenotype <- inner_join (s1_features_with_outcome, s1_phenotype, by="PATIENTID")
```

do phenotypes correlate with outcome?
```{r}
chisq.test(table(s1_features_with_outcome_phenotype$SEX, s1_features_with_outcome_phenotype$SURVIVAL_STATUS))

chisq.test(table(s1_features_with_outcome_phenotype$RACE, s1_features_with_outcome_phenotype$SURVIVAL_STATUS))

chisq.test(table(s1_features_with_outcome_phenotype$WHO_GRADING, s1_features_with_outcome_phenotype$SURVIVAL_STATUS))

chisq.test(table(s1_features_with_outcome_phenotype$CANCER_TYPE, s1_features_with_outcome_phenotype$SURVIVAL_STATUS))

```

```{r}

s1_features_with_outcome_phenotype %>% 
  group_by(CANCER_TYPE, SURVIVAL_STATUS) %>%
  summarise(num_patient = n()) %>%                     
  ggplot(aes(x=CANCER_TYPE, y=num_patient, fill=as.factor(SURVIVAL_STATUS))) +
  geom_bar(stat="identity", position="fill")
```

all unclassified cases are dead, how many are unclassified?
```{r}
s1_features_with_outcome_phenotype %>% 
  group_by(CANCER_TYPE, SURVIVAL_STATUS) %>%
  summarise(num_patient = n()) 

```

drop patientID
```{r}
s1_rna_features <- s1_rna_features %>% select(-PATIENTID)
s1_features_with_outcome <- s1_features_with_outcome %>% select(-PATIENTID)
```

one hot encoding 
# and drop patientID, RACE and GRADING
```{r}
dmy <- dummyVars(" ~ CANCER_TYPE", data = s1_features_with_outcome_phenotype)
CANCER_TYPE_encodings <- data.frame(predict(dmy, newdata = s1_features_with_outcome_phenotype))
s1_features_with_outcome_phenotype <- cbind(s1_features_with_outcome_phenotype, CANCER_TYPE_encodings)

#s1_features_with_outcome_phenotype <- s1_features_with_outcome_phenotype %>% select(-PATIENTID, -RACE,  -WHO_GRADING, -CANCER_TYPE )
```

save these
```{r}
saveRDS(s1_rna_features, file = here("munge", "s1_rna_features.rds"))
saveRDS(s1_features_with_outcome, file = here("munge", "s1_features_with_outcome.rds"))
saveRDS(s1_features_with_outcome_phenotype, file = here("munge", "s1_features_with_outcome_phenotype.rds"))
```