---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

Here I will use the models trained in s1 and s2 for the prediction model in s3


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(caret)
library(pROC)
library(h2o)
h2o.init()

seed <- 17
```

load data and models
```{r}
s3_features_with_outcome_phenotype <- readRDS(file = here("munge", "s3_features_with_outcome_phenotype.rds"))

rna_predictors.all.result <- scan(here("/munge/rna_borutaAllFeatures.txt"), character(), quote = "") 
rna_predictors.overlap.result <- scan(here("/munge/rna_borutaOverlapFeatures.txt"), character(), quote = "") 

cn_predictors.all.result <- scan(here("/munge/cn_borutaAllFeatures.txt"), character(), quote = "") 
cn_predictors.overlap.result <- scan(here("/munge/cn_borutaOverlapFeatures.txt"), character(), quote = "") 

rna_model <- h2o.loadModel(path = here("/results/h2o_automl_rna_model//DeepLearning_grid__1_AutoML_20200122_013655_model_1"))
cn_model <-  h2o.loadModel(path = here("/results/h2o_automl_cn_model/DeepLearning_grid__1_AutoML_20200122_041609_model_1"))
  
s1_test.data_with_outcome_phenotype <- readRDS(file = here("munge", "s1_test.data_with_outcome_phenotype.rds"))
s2_test.data_with_outcome_phenotype <- readRDS(file = here("munge", "s2_test.data_with_outcome_phenotype.rds"))
```

test data as intersect and union of the test sets before
it seems that the intersect has too few samples, we will test on the union
```{r}

test_patientIDs_intersect <- intersect (s1_test.data_with_outcome_phenotype$PATIENTID, s2_test.data_with_outcome_phenotype$PATIENTID)

test_patientIDs_union <- union (s1_test.data_with_outcome_phenotype$PATIENTID, s2_test.data_with_outcome_phenotype$PATIENTID)

length(test_patientIDs_intersect)
length(test_patientIDs_union)

```

Generate test set:
on hot encoding
change SURVIVAL_STATUS to class
```{r}
testset_s3 <- s3_features_with_outcome_phenotype %>% filter (PATIENTID %in% test_patientIDs_union)
table(testset_s3$SURVIVAL_STATUS)


testset_s3 <- testset_s3 %>% 
                            mutate (SURVIVAL_STATUS = as.factor (SURVIVAL_STATUS)) %>%
                            rename (class = SURVIVAL_STATUS)

```


accuracy of rna model: 0.88
accuracy of cn model: 0.6176
take a weighted prediction from the two models
```{r}
testset_s3_h2o <- as.h2o(testset_s3, destination_frame="testset_s3_h2o")

testset_s3_rna_pred <- as.data.frame(h2o.predict(rna_model, testset_s3_h2o))
testset_s3_cn_pred <- as.data.frame(h2o.predict(cn_model, testset_s3_h2o))

testset_s3_rna_cn_pred <- data.frame( p1=(testset_s3_rna_pred$p1*0.88 + testset_s3_cn_pred$p1*0.6176)/(0.88+0.6176) )  %>%
                          mutate(predict = if_else (p1>=0.5, 1, 0))

```

confusion matrix
```{r}


confusionMatrix(as.factor(testset_s3_rna_cn_pred$predict), 
                as.factor(testset_s3$class), positive = "1")

```

AUC
```{r}
s3_roc <- roc(as.factor(testset_s3$class), testset_s3_rna_cn_pred$p1)
auc(s3_roc)

```